---
title: "Strings and factors"
author: "Diana Hernandez"
date: "2023-10-17"
output: github_document
---

```{r}
library(tidyverse)
library(rvest)
library(p8105.datasets)
```

# Strings and manipulations

`str_detect` finds the pattern within the string. Has to be exact match.

`str_replace` can replace one string for another string.

```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")

str_detect(string_vec, "Jeff")

str_replace(string_vec, "jeff", "Jeff")
```

start with: ^
ends with: $

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think")

str_detect(string_vec, "^i think")

str_detect(string_vec, "i think$")
```

If you want both lower and uppercase: [Pp]

```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"pumpkin")
str_detect(string_vec,"Pumpkin")
str_detect(string_vec,"[Pp]umpkin")
```

Can detect multiple strings

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9]")
str_detect(string_vec, "^[0-9]")
str_detect(string_vec, "^[0-9][a-z]")
str_detect(string_vec, "^[0-9][a-zA-Z]")
```

The character `.` matches anything.

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```

Some characters are “special”. These include `[` and `]`, `(` and `)`, and .. If you want to search for these, you have to indicate they’re special using `\`. Unfortunately, `\` is also special, so things get weird. Two `\\` to find a special character that is not `\`.

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```

# Factors

`fct_relevel` to relevel a factor. Changes the order.

```{r}
vec_sex = factor(c("male", "male", "female", "female"))

vec_sex

as.numeric(vec_sex)

vec_sex = fct_relevel(vec_sex, "male")

vec_sex

as.numeric(vec_sex)
```

# NSDUH

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) |> 
  html_table() |> 
  first() |>
  slice(-1)
```

need to tidy this dataset!

```{r}
marj_df =
  table_marj |> 
  select(-contains("P Value")) |> 
  pivot_longer(
    !State,
    names_to = "age_year",
    values_to = "percent"
  ) |> 
  separate(age_year, into = c("age", "year"), "\\(") |> 
  mutate(year = str_replace(year, "\\)", ""),
         percent = str_replace(percent, "[a-b]", ""),
         percent = as.numeric(percent)
  ) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```


```{r}
marj_df |>
  filter( age == "18-25") |>
  mutate(State = fct_reorder(State, percent)) |>
  ggplot(aes(x = State, y = percent, color = year)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


